SAT Solving for Vesicle Traffic Systems in Cells

### 2017-07-20 paper/encoding.tex
wd

\subsubsection{$k$-connectivity constraints}
To check whether $k$-connected is a necessary condition, we remove (drop) $k-1$ edges from the graph and if it
disconnects the graph, and we get an assignment. We have a graph that is not  $k$-connected.
%
Boolean constraints~\ref{eq:drop1}-\ref{eq:drop4}
encode the relevant constraints for reachability
in the modified VTS. 
%
The following constraints encode that only
existing edges can be dropped and exactly $k-1$ edges are dropped.
\begin{align}
  \bigwedge\limits_{i,j,q} d_{i,j,q} \limplies e_{i,j,q}  \tag{D1}\label{eq:drop1}\\
  \sum_{i,j,q} d_{i,j,q} = k-1
  \tag{D2}\label{eq:drop2}
\end{align}
We need to define reachability in the modified VTS, therefore we use
a new variable $r'_{i,j}$ to encode reachability from $i$ to $j$.
In the following constraint, we encode $r'_{i,j}$ is true if there is an
edge $(i,\_,j)$ and it is not dropped, or there is a node
$i^{\prime}$ such that, there is an edge $(i,\_,i^{\prime})$ which is
not dropped and $r'_{i',j}$ is true.
\begin{align}
\bigwedge\limits_{i,j}  [\bigvee_{q} (e_{i,j,q} \land  \neg d_{i,j,q}) \lor  (\bigvee_{i' \neq i}  r^{\prime}_{i',j} \land  \bigvee_{q} (e_{i,i',q} \land \neg d_{i,i',q}) ] \limplies r^{\prime}_{i,j}  
  \tag{D3}\label{eq:drop3}
\end{align}
Since we are looking for a graph that is not $k$-connected,
the following constraint encodes that there are nodes $i$ and $j$ such that
there is no path between them in the underlying undirected graph.
\begin{align}
   \bigvee\limits_{i,j} \neg (r^{\prime}_{i,j} \lor r^{\prime}_{j,i})
  \tag{D4}\label{eq:drop4}
\end{align}

>>>>

\subsubsection{$k$-connectivity constraints}
To check whether $k$-connected is a necessary condition, we remove (drop) $k-1$ edges from the graph and if it
disconnects the graph, and we get an assignment. We have a graph that is not  $k$-connected.
%
Constraints~\ref{eq:drop1}-\ref{eq:drop4}
encode the relevant constraints for reachability
in the modified VTS. 
%
The following constraints encode that only
existing edges can be dropped and exactly $k-1$ edges are dropped.
\begin{align}
  \bigwedge\limits_{i,j,q} d_{i,j,q} \limplies e_{i,j,q}  \tag{D1}\label{eq:drop1}\\
  \sum_{i,j,q} d_{i,j,q} = k-1
  \tag{D2}\label{eq:drop2}
\end{align}
We need to define reachability in the modified VTS, therefore we use
a new variable $r'_{i,j}$ to encode reachability from $i$ to $j$.
In the following constraint, we encode $r'_{i,j}$ is true if there is an
edge $(i,\_,j)$ and it is not dropped, or there is a node
$i^{\prime}$ such that, there is an edge $(i,\_,i^{\prime})$ which is
not dropped and $r'_{i',j}$ is true.
\begin{align}
\bigwedge\limits_{i,j}  [\bigvee_{q} (e_{i,j,q} \land  \neg d_{i,j,q}) \lor  (\bigvee_{i' \neq i}  r^{\prime}_{i',j} \land  \bigvee_{q} (e_{i,i',q} \land \neg d_{i,i',q}) ] \limplies r^{\prime}_{i,j}  
  \tag{D3}\label{eq:drop3}
\end{align}
Since we search for disconnected modified VTS,
the following constraint encodes that there are 
unreachable pair of nodes in the underlying undirected graph.
% nodes such that
% there is no path between them .
\begin{align}
   \bigvee\limits_{i,j} \neg (r^{\prime}_{i,j} \lor r^{\prime}_{j,i})
  \tag{D4}\label{eq:drop4}
\end{align}

### 2017-07-20 paper/experiments.tex
wd

% C_es :Self edges are allowed
% C_ed : Every edge is distinct 
  \end{tabular}
\caption{{\bf Activity regulation of molecules vs. graph connectivity.}}
\label{tab:grph}
\end{table}

@@

We have implemented the encodings for each variants
using the python interface of $\zthree$ in a tool (A'tool).
%
Our tool allows the user to choose a model and the size of the network
besides other parameters like connectivity and number of parallel
edges.
%
It uses Z3 Python interface to build the needed constraints and
applies $\zthree$ solver on the constraints to find a model (
a satisfying assignment that respects the constraints).
%
This tool also translates the satisfying model found by $\zthree$ into
a VTS and presents a visual output to the user in form of annotated graph.
%
% The output graph will satisfy the underlying rules of the system, so
% it always is a valid vesicle transport network.
%
% The resultant graph provides an overview of the underlying VTS
% where directed labeled edges and nodes provide the
% complete information about the system.
%
We also visually report the dropped edges required to disconnect the graph, it gives information about the connectivity of the graph.

@@

In Table~\ref{tab:qf-grabh}, we present the running times for the search of
VTSs of sizes 2 to 10 that satisfy the variants.
%
The experiments were done on a machine with Intel(R) Core(TM) i3-4030U
CPU @ 1.90GHz processor and 4GB RAM.
%
We have compared our performance with the performance of our earlier 
CBMC based implementation (O'cbmc).
%
For example, the formula for variation F, the total number of
compartments ($|N|$) equals to 10, returns in 129.78 minutes (7786.8 secs)
with a SAT result.
%
In comparison, CBMC results in OUT OF MEMORY for $|N|$ greater than 5.
%
``!'' indicate that the constraints were unsatisfiable.
%
 \textbf{Using this encoding in comparison to the old one, not only we got efficiency improved for finding a SAT model but also did better in the case of refutation that no model exist (Table 2 variant A timing comparison and variant D with N =2).}. Hence with the use of this novel encoding, we are able to scale the system to a much larger compartmentalized system, especially to
eukaryotic cells with a total number of 10 compartments.
%
% \ankit{
% This additional scalability can provide us leverage to ask for related
% open questions that were not previously possible, for example, what is
% the minimum number of molecules required to satisfy the vesicular
% traffic system in eukaryotes.}
%
Furthermore, we experimented with limits of our tools and found
that $\zthree$ was able to solve the constraints up to $\sim{14-18}$ nodes.

>>>>

% C_es :Self edges are allowed
% C_ed : Every edge is distinct 
  \end{tabular}
\caption{{\bf Activity regulation of molecules vs. graph connectivity.}}
\label{tab:grph}
\vspace{-4ex}
\end{table}

@@

We have implemented~\footnote{\url{https://github.com/arey0pushpa/pyZ3}}
the encodings for each variants
using the python interface of $\zthree$ in a tool (A'tool).
%
Our tool allows the user to choose a model and the size of the network
besides other parameters like connectivity and number of parallel
edges.
%
It uses Z3 Python interface to build the needed constraints and
applies $\zthree$ solver on the constraints to find a model (
a satisfying assignment that respects the constraints).
%
This tool also translates the satisfying model found by $\zthree$ into
a VTS and presents a visual output to the user in form of annotated graph.
%
% The output graph will satisfy the underlying rules of the system, so
% it always is a valid vesicle transport network.
%
% The resultant graph provides an overview of the underlying VTS
% where directed labeled edges and nodes provide the
% complete information about the system.
%
We also visually report the dropped edges required to disconnect the graph, it gives information about the connectivity of the graph.

@@

In Table~\ref{tab:qf-grabh}, we present the running times for the search of
VTSs of sizes 2 to 10 that satisfy the variants.
%
The experiments were done on a machine with Intel(R) Core(TM) i3-4030U
CPU @ 1.90GHz processor and 4GB RAM.
%
We have compared our performance with the performance of our earlier 
CBMC based implementation (O'cbmc).
%
For example, the formula for variation F, the total number of
compartments ($|N|$) equals to 10, returns in 129.78 minutes (7786.8 secs)
with a SAT result.
%
In comparison, CBMC results in OUT OF MEMORY for $|N|$ greater than 5.
%
``!'' indicate that the constraints were unsatisfiable.
%
Using this encoding in comparison to the old one, not only we got
efficiency improved for finding a SAT model but also did better in the
case of refutation that no model exist (Table 2 variant A timing
comparison and variant D with N =2).
%
Hence with the use of this novel encoding, we are able to scale the system to a much larger compartmentalized system, especially to
eukaryotic cells with a total number of 10 compartments.
%
% \ankit{
% This additional scalability can provide us leverage to ask for related
% open questions that were not previously possible, for example, what is
% the minimum number of molecules required to satisfy the vesicular
% traffic system in eukaryotes.}
%
Furthermore, we experimented with limits of our tools and found
that $\zthree$ was able to solve the constraints up to $\sim{14-18}$ nodes.
